---
title: "Meta-analysis of proportions"
author: "Valirie N. Agbor"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

# Remove objects in the working directory --
rm(list = ls())

# Load packages--
pacman::p_load(
  tidyverse, 
  stringr, 
  readxl,
  janitor,
  meta, metafor
)

knitr::opts_knit$set(root.dir = "./temp")
```


# Load data --

```{r Read data}

# Get datasets in directory--
list_files <- list.files(getwd(), pattern = ".xlsx", recursive = TRUE)

# Get file names to use as forest plot labels --
name_file <- list_files %>% 
                str_replace(pattern = ".xlsx", "") %>% 
                str_replace(pattern = "Datasets/", "")


# Create list of datasets -- 

## Loop through file names with purrr
## Read datasets and hold the in a list
## Use purrr to loop through datasets and do some cleaning 

dat_list <- 
  
  purrr::map2(
   .x = list_files,     
   .y = seq_along(name_file), 
      ~ readxl::read_excel(path = paste0("./", .x)) %>% 
        janitor::clean_names()  %>% 
            mutate(
               study       = paste(author, year, sep = ", "),  
               unsd_region = str_to_title(unsd_region),
               year_group  = factor(case_when(year <= 2010 ~ "2010 or earlier", TRUE ~ "After 2010")),
               name        = name_file[.y]
               )
   )
   
```


# Run meta-analysis --

Fit a random-effect meta-analysis using a DL estimator. 
Stabilise the variance of study-specific estimates using the Freeman-Tukey double-arcsine transformation 


## Diagnostics with metafor --

```{r Metafor - Summary prop}

# Transform proportions

ies <- purrr::map(.x = dat_list, 
           ~metafor::escalc(xi      = cases,
                   ni      = sample_size,
                   measure = "PFT", 
                   data    = .x, 
                   add     = 0))


pes.da <- purrr::map(.x = ies, 
              ~metafor::rma(
                   yi, 
                   vi, 
                   data     = .x, 
                   method   = "DL", 
                   weighted = TRUE, 
                   slab     = study))


pes <- purrr::map2(.x = pes.da, .y = dat_list,
                  ~metafor::predict(
                    .x,
                    transf = transf.ipft.hm,
                    targ   = list(ni = .y$sample_size))
           )   


knitr::opts_chunk$set(echo = FALSE)

```

### Identify outliers -- 

```{r Baujat plot}

baujat_list <- purrr::map(.x = pes.da, ~baujat(.x, symbol = "slab")) 

```

```{r Influential studies}

inf <- purrr::map(.x = pes.da, ~influence(.x))

purrr::map(.x = inf, ~plot(.x))

```

```{r leave-1-out}

l1o <- purrr::map(.x = pes.da, ~metafor::leave1out(.x))

yi <- purrr::map(.x = l1o, ~.x$estimate)
vi <- purrr::map(.x = l1o, ~.x$se^2)

purrr::map2(
     .x = yi, 
     .y = seq_along(yi), 
      ~metafor::forest(
             .x,
             vi[[.y]],
             transf = transf.ipft.hm,
             targ   = list(ni = dat_list[[.y]]$sample_size),
             slab   = dat_list[[.y]]$study,
             xlab   = "Summary proportions leaving out each study",
             refline= pes[[.y]]$pred,
             digits = 6)
     )

```


## Forest plot -- 

Once you are satisfied that there are no influential studies, fit the meta-analysis and make forest plots

```{r forestplot, fig.height = 11, fig.width = 11}


MA_random <- 
  purrr::map(.x = dat_list, 
       ~meta::metaprop(data        = .x, 
                  event       = cases, 
                  n           = sample_size,
                  studlab     = study,  
                  sm          = "PFT", 
                  method.ci   = "CP", 
                  level       = 0.95, 
                  method.tau  = "DL", 
                  pscale      = 100,
                  digits.pval = 4, 
                  prediction  = TRUE, 
                  level.ma    = 0.95, 
                  random      = TRUE, 
                  fixed       = FALSE
                  )
             )


# Make forestplot ----

for (a in seq_along(MA_random)) {

  # Get length of dataset
  n_row <- nrow(dat_list[[a]])
     
  png(paste0(name_file[[a]], ".png"), 
      units     = "in",
      width     = 10, 
      height    = ifelse(n_row <5, 5, 
                         ifelse(n_row >= 5 & n_row <10, 6, 
                                ifelse(n_row >= 10, 7, 6))), # Programmatically set page height
      pointsize = 15,  
      res       = 300, 
      bg        = "white") 

  
  #Forest plot
  meta::forest(MA_random[[a]],
         xlim        = c(0, 100), 
         pscale      = 100,
         rightcols   = c("effect", "ci","w.random"),
         rightlabs   = c("Prevalence",  "95% CI","    Weight(%)"),
         leftcols    = c("studlab", "event", "n"), 
         leftlabs    = c("Study", "          Cases", "  Sample"),
         xlab        = "Prevalence (%)",
         xlab.pos    = 50,
         ff.xlab     = "bold", 
         fs.axis     = 11,
         sortvar     = study,
         fs.lr       = 10,
         fs.xlab     = 13,
         fs.study    = 10.3,
         fs.study.lables=10.3,
         fs.heading=11,
         fs.random = 10.3, 
         fs.predict = 10.3,
         fs.predict.labels = 10.3,
         fs.hetstat=10,
         squaresize = 0.45, 
         col.square="gray", 
         col.square.lines="gray",
         col.diamond="black", 
         col.diamond.lines="black",
         comb.fixed=FALSE,
         comb.random=FALSE,
         lty.fixed=0,
         lty.random=2, 
         type.study="square",
         type.random="diamond",
         #ff.fixed="bold.italic",
         ff.random="bold",
         hetlab = "Heterogeneity:",
         spacing = 1.3,
         smlab="",
         plotwidth = "6cm",
         pooled.events = TRUE,
         print.Q=FALSE,
         print.tau2.ci = FALSE,
         print.pval.Q=TRUE,
         print.I2=TRUE,
         print.tau2=TRUE,
         prediction = TRUE,
         col.by="grey",
         digits.addcols.right = 1,
         digits.I2 = 1, 
         digits.pval.Q = 4, 
         digits.weight = 1, 
         digits.tau2 = 4,
         digits = 1,
         fixed = FALSE)


dev.off()

}


```


## Subgroup --

Only do subgroup analysis if number of studies >= 10

```{r forestplot, fig.height = 11, fig.width = 11}

# Get data for subgroup analysis 
subgroup_vars <- c("age_group", "year_group", "unsd_region")
subgroup_name <- c("Age group", "Publication Year", "UNSD region")


subgroup_data <- 
  
    purrr::keep(.x = dat_list, ~nrow(.x) >= 10) %>% 
    purrr::map( 
     ~dplyr::select(
      # Select relevant variables for subgroup analysis --
        .x, 
        study, 
        cases, 
        sample_size,
        all_of(subgroup_vars),
        name
        )
     )


# Run subgroup analysis ---
result_subgroup <- list()

for (df in seq_along(subgroup_data)) { # Loop through datasets for subgroup analysis

  subgroup_temp <- list()
  df_temp <- subgroup_data[[df]]

    
    # Loop over subgroup vars --
    for (s in seq_along(subgroup_vars)) { # Loop through subgroup variables --
    
    sub_var_temp      <- subgroup_vars[s]
    sub_var_lab_temp  <- subgroup_name[s]
    
    
    subgrp_temp <- 
      
   meta::metaprop(data          = df_temp, 
                  cases, 
                  sample_size,
                  studlab       = df_temp[["study"]],  
                  sm            = "PFT", 
                  method.ci     = "CP", 
                  level         = 0.95, 
                  method.tau    = "DL", 
                  pscale        = 100,
                  digits.pval   = 4, 
                  prediction    = TRUE, 
                  level.ma      = 0.95, 
                  random        = TRUE, 
                  fixed         = FALSE,
                  subgroup      = df_temp[[sub_var_temp]], 
                  subgroup.name = sub_var_lab_temp
                  ) 
   
    subgroup_temp[[sub_var_temp]] <- 
           update(
            subgrp_temp, 
            byvar = subgrp_temp$data[[".subgroup"]], 
            subgroup.name = sub_var_lab_temp)
    
    # End of Loop for subgroup variables --
    
    }
  
  # Bind results of each subgroup --
  
  mbind <- meta::metabind(subgroup_temp[[1]], 
                          subgroup_temp[[2]], 
                          subgroup_temp[[3]]) # The index depends on the no. of subgroups --
  
  result_subgroup[[df]] <- mbind
  
}


# Make summary forest plot for subgroup analysis --

for (p in seq_along(result_subgroup)) { 

  png(paste0(unique(subgroup_data[[p]]$name), ".png"), 
             units     = "in", 
             width     = 11, 
             height    = 8, 
             pointsize = 12,  
             res       = 600)  

  meta::forest(result_subgroup[[p]], 
             spacing     = 1.3, 
             plot.width  = "6cm",
             xlab        = "\nPrevalence (%)",
             xlab.pos    = 50,
             xlim        = c(0, 100),
             rightcols   = c("effect", "ci", "I2"), 
             rightlabs   = c("Prevalence", "95% CI", "Het"),
             leftlabs    = c("Studies (N)", "Interaction\np-value"),
             smlab       = "Random effects",
             colgap      = "4mm",
             squaresize  = 0.5, 
             col.by      = "dark blue", 
             print.subgroup.labels = TRUE)
      
  dev.off()
  
                         
    }


```



## Funnel plots and publication bias --

```{r Funnel plots}

## Get p-values for publication bias
df_pub_bias <- keep(.x = MA_random, ~.x$k>=5) # Keep dataset with at least 5 rows


# Extract p-values --
bias_p <- 
  map(
    .x = df_pub_bias, 
    ~metabias(.x, k.min = 5, method.bias = "linreg")) %>% map_dbl(~round(.x$pval, 4))


# Make funnel plots --

for (b in seq_along(df_pub_bias)) {

png(paste0(unique(df_pub_bias[[b]]$data$name),".png"), 
    units     = "in", 
    width     = 10, 
    height    = 7, 
    res       = 600, 
    pointsize = 12)

    meta:::funnel.meta(df_pub_bias[[b]],
                backtransf = TRUE,
                col     = "navy blue",
                back    = "white",
                xlab    = "Freeman-Tukey double arcsine transformed proportion")
    
    text(x = 0.4, y = 0.01, label = paste0("p(Egger)=", bias_p[[p]]))

dev.off()

}
 

```

